### 2021-05-12 report/paper.tex
Add template for group report

\documentclass[]{usiinfprospectus}

\captionsetup{labelfont={bf}}

\author{Bevilacqua Joey \\[3pt] De Vita Gianmarco  \\[3pt] Luini Alessandro  \\[3pt] Maggioni Claudio}

\title{\centerline{Sudoku}}
\versiondate{\today}

\abstract {
Abstract}

\begin{document}
\maketitle
\tableofcontents
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{introduction}

%%%%%
\bibliographystyle{abbrv}
\bibliography{references}

\end{document}