### 2021-05-23 report/paper.tex
Add problem definition, report updat

\documentclass[]{usiinfprospectus}

\captionsetup{labelfont={bf}}

\author{\centerline{Bevilacqua Joey} \\[3pt] \centerline{De Vita Gianmarco}  \\[3pt] \centerline{Luini Alessandro}  \\[3pt] \centerline{Maggioni Claudio}}

\title{\centerline{Sudoku}}
\versiondate{\today}
\usepackage{boldline} 
\usepackage{enumitem}

\abstract {
The aim of this project is to make an Encoder and Decoder for a SAT-solver in order to decide whether a problem is satisfiable or unsatisfiable. WRITE A BETTER ABSTRACT
}

% For Sudoku drawing
\usepackage{tikz}
\newcounter{row}
\newcounter{col}

\newcommand\setrow[9]{
  \setcounter{col}{1}
  \foreach \n in {#1, #2, #3, #4, #5, #6, #7, #8, #9} {
    \edef\x{\value{col} - 0.5}
    \edef\y{9.5 - \value{row}}
    \node[anchor=center] at (\x, \y) {\n};
    \stepcounter{col}
  }
  \stepcounter{row}
}

\begin{document}
\maketitle
\tableofcontents
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} \label{introduction}
Even though Sudoku is a Japanese word, literally meaning "digit-single", according to some sources it was actually invented in Switzerland by the already famous mathematician Euler from Basel. However, it is generally acknowledged that the origin of this game is European (around XIX century) and that the modern and common version has been realized in the United States in the second half of the XX century named as "Number Place", from which it spread across the globe, even in Japan, where it took the famous name it is known with.\\ \\
The aim of this homework is to make an Encoder and Decoder for a SAT-solver in order to decide whether a problem is satisfiable or unsatisfiable. 
\newpage
\section{Problem definition} \label{problem}
Sudokus are puzzles based upon number placement. They are classically structured in a $9\times 9$ grid with 81 cells, divided into $3\times 3$ boxes. In order to be satisfiable, the following statements must hold:
\begin{itemize}%[label={(\arabic*)}]
\item In every cell there is exactly one number $c \in \{1, 2, ..., 9 \}$
\item No number may appear more than once in the same row or column.
\item No number may appear more than once in the same box.
\end{itemize}
Often Sudoku puzzles are initialized with some values placed, leaving blank spaces to be filled by the player with values respecting the rules above. Visually, Sudokus appear as a square grid, an example can be the following:

%Draw Sudoku
\begin{figure}[h]
\begin{center}
\begin{tikzpicture}[scale=.7]
  \begin{scope}
    \draw (0, 0) grid (9, 9);
    \draw[very thick, scale=3] (0, 0) grid (3, 3);

    \setcounter{row}{1}
    \foreach \i in {1,...,9}
    {
    \setrow {$c_{\i,1}$ }{$c_{\i,2}$}{$c_{\i,3}$}  {$c_{\i,4}$}{$c_{\i,5}$}{$c_{\i,6}$}  {$c_{\i,7}$}{$c_{\i,8}$}{$c_{\i,9}$}
    }
  \end{scope}
\end{tikzpicture}
\end{center}
\caption{A $9\times 9$ Sudoku puzzle.}
\end{figure}
\noindent
Assuming that every $c_{i,j}$ in the Sudoku grid in Figure 1 takes just a value in interval \{1,2,...,9\}, such solution can be considered valid if and only if the following conditions are satisfied: 
\begin{enumerate}[label={(\arabic*)}]
\item In every row $i$ (i.e., $c_{i,1}$, $c_{i,2}$, ..., $c_{i,9}$) there is at least one element that takes one value in \{1,2,...,9\}, but all values from 1 to 9 must be taken. Since in a row there are only 9 elements, it means that there will be a one-to-one mapping from these row elements to the \{1,2,...,9\} set values, thus each row will be made of nine different numbers.
\item In every column $j$ (i.e., $c_{1,j}$, $c_{2,j}$, ..., $c_{9,j}$) there is at least one element that takes one value in \{1,2,...,9\}, but all values from 1 to 9 must be taken. Since in a column there are only 9 elements, it means that there will be a one-to-one mapping from these column elements to the \{1,2,...,9\} set values, thus each column will be made of nine different numbers.
\item In every box, no number may appear more than once. Since a box is made up by 9 elements, it means that each box will contain all the values within the \{1,2,...,9\} set. What requires a bit more of reasoning here is finding a generalizing way to define all the elements in a box for every box. While for rows and columns it suffices to use a matrix-like notation, here some background reasoning is needed. It can be noticed that, in every box, the following pattern can be observed:
\begin{figure}[h]
\begin{center}
\begin{tikzpicture}[scale=.7]
  	\begin{scope}
    	\draw (0, 0) grid (3, 3);
   	\draw[very thick, scale=3] (0, 0) grid (1, 1);

   	\node[anchor=center] at (0.5, 3.5) {\scriptsize{$j$}};
   	\node[anchor=center] at (1.5, 3.5) {\scriptsize{$j+1$}};
   	\node[anchor=center] at (2.5, 3.5) {\scriptsize{$j+2$}};

   	\node[anchor=center] at (-0.5, 2.5) {\scriptsize{$i$}};
   	\node[anchor=center] at (-0.5, 1.5) {\scriptsize{$i+1$}};
   	\node[anchor=center] at (-0.5, 0.5) {\scriptsize{$i+2$}};

   	\node[anchor=center] at (0.5, 2.5) {$c_1$};
   	\node[anchor=center] at (1.5, 2.5) {$c_2$};
   	\node[anchor=center] at (2.5, 2.5) {$c_3$};

   	\node[anchor=center] at (0.5, 1.5) {$c_4$};
   	\node[anchor=center] at (1.5, 1.5) {$c_5$};
   	\node[anchor=center] at (2.5, 1.5) {$c_6$};

   	\node[anchor=center] at (0.5, 0.5) {$c_7$};
   	\node[anchor=center] at (1.5, 0.5) {$c_8$};
   	\node[anchor=center] at (2.5, 0.5) {$c_9$};

  	\end{scope}
\end{tikzpicture}
\end{center}
\caption{A $3\times 3$ Sudoku generalized box.}
\end{figure}

\noindent
That means that every box has its top-leftmost cell at a particular coordinate ($i,\,j$) and its other corner cells are clockwise respectively at coordinates: ($i,\,j+2$),  ($i+2,\,j+2$) and ($i+2,\,j$). By this, it can be said that by individuating the top-leftmost cell of a box it is possible to individuate the box itself. Hence, for $i=1$ and $j=1$, we have corner cells ($1,\,1$), ($1,\,3$),  ($3,\,3$) and ($3,\,1$), that is the top-leftmost box in the Sudoku grid. From this approach, the boxes making up the grid can be defined as:
\begin{figure}[h]
\begin{center}
\begin{tikzpicture}[scale=1]
  	\begin{scope}
    	\draw (0, 0) grid (3, 3);
   	\draw[very thick, scale=3] (0, 0) grid (1, 1);

   	\node[anchor=center] at (0.5, 2.5) {$(1,1)$};
   	\node[anchor=center] at (1.5, 2.5) {$(1,4)$};
   	\node[anchor=center] at (2.5, 2.5) {$(1,7)$};

   	\node[anchor=center] at (0.5, 1.5) {$(4,1)$};
   	\node[anchor=center] at (1.5, 1.5) {$(4,4)$};
   	\node[anchor=center] at (2.5, 1.5) {$(4,7)$};

   	\node[anchor=center] at (0.5, 0.5) {$(7,1)$};
   	\node[anchor=center] at (1.5, 0.5) {$(7,4)$};
   	\node[anchor=center] at (2.5, 0.5) {$(7,7)$};

  	\end{scope}
\end{tikzpicture}
\end{center}
\caption{A $9\times 9$ Sudoku grid. Each square corresponds to a $3\times 3$ box. Inside of the square, the ($i,\,j$) coordinate of the top-leftmost cell of that box is written.}
\end{figure}

\noindent
It can be seen that the coordinates are nothing but all the possible permutations of three values: 1, 4 and 7. From this, we can define a set $B=\{1,\,4,\,7\}$ containing these values.\\
\end{enumerate}
\noindent
Let \textbf{isvalid} define whether a row, column or box is valid. That means that the nine elements along a row, column or within a box must be different and equal to a number in set $\{1,2,...,9\}$.
\begin{align}
\text{isvalid}\left(  c_1,  c_2,  c_3,  c_4,  c_5,  c_6,  c_7,  c_8,  c_9  \right) = \bigwedge^9_{i=1}  \bigvee^9_{i=1} \,\,\, c_i
\end{align}

\noindent
Therefore, the Sudoku problem can be defined as:
\setcounter{equation}{0}
\begin{align}
\text{sudoku}\left( \{ c_{i,j} \}_{i, j \in \{ 1,...,9\}} \right) = & \nonumber\\
&\bigwedge^9_{i=1} \text{ isvalid}\left(  c_{i,1},  c_{i,2},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9}  \right)\\
\wedge &\bigwedge^9_{j=1} \text{ isvalid}\left( c_{1,j},  c_{2,j},  c_{3,j},  c_{4,j},  c_{5,j},  c_{6,j},  c_{7,j},  c_{8,j},  c_{9,j} \right) \\
\wedge &\bigwedge^9_{i,j \in B} \text{ isvalid}\left( c_{i,1},  c_{i,(j+1)},  c_{i,3},  c_{i,4},  c_{i,5},  c_{i,6},  c_{i,7},  c_{i,8},  c_{i,9} \right).
\end{align}
\section{Problem definition} \label{problem}
\section{Problem definition} \label{problem}
%%%%%
\bibliographystyle{abbrv}
\bibliography{references}

\end{document}